### 2017-06-01 m.Rnw

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(warning=F, message = FALSE,echo=F,results='asis',error=FALSE,fig.lp='fig:',fig.path='images/')
@

<<mortality,fig.cap='Mortality as a function of activity $\\tau$ for three levels of the mortality scaler $\\eta$',fig.width=5, fig.height=4,>>=
  require(ggplot2)
mframe <- data.frame(tau = rep(c(0,exp(seq(log(0.001),log(1),l=100))),each=3),
                     etalab = as.character(c(0.33,1,3)),
                     M=rep(c(0,exp(seq(log(0.001),log(1),l=100))),each=3)^c(0.33,1,3)*1)

ggplot(mframe) + 
  geom_line(aes(x=tau,y=M,linetype=etalab)) + 
  theme_bw() +
  xlab(bquote(tau)) +
  scale_linetype(expression(eta))
@