### 2017-06-30 figures/O2_fig.tex
arg

\begin{figure}
\includegraphics[width=\linewidth]{images/O2_fig-1} \caption[Maximum metabolic rate, determined by oxygen supply, as a function of temperature, for four different optimal temperatures (panels---see panel label for optimal temperature in degrees celsius), and increasing values of $\eta$]{Maximum metabolic rate, determined by oxygen supply, as a function of temperature, for four different optimal temperatures (panels---see panel label for optimal temperature in degrees celsius), and increasing values of $\eta$.}\label{fig:O2_fig}
\end{figure}