### 2017-08-18 poster/poster.Rmd
commit before changing poster to norm-based

---
title: "Sizing the effects of temperature on fish"

author: ['<span class=\"presenter\">Philipp Neubauer<sup>1</sup></span>, Ken H. Andersen<sup>2</sup><ol class=\"affiliations\"><li> Dragonfly Data Science, Wellington, New Zealand</li><li>Centre for Ocean Life, National Institute of Aquatic Resources, Technical University of Denmark, Denmark</li></ol>']

output: drposter::revealjs_poster

bibliography: ../Sizing-fx-of-temperature.bib
csl: nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = F,message = F,
                      cache=T,autodep = T,dev = "svg")  # Hide code by default. Can override chunk-by-chunk
```

# {.col-1}

## Introduction

Size-at-maturation and maximum size are key traits, defining ecoloical niches and driving ecological interactions [@andersen_asymptotic_2006]. Temperature has been identified in various contexts as a key driver of organism size (Bergman's rule, Temperature size rule...) [@atkinson_temperature_1994]. Climate change may therefore have strong effects on organism size and follow on effects at population and ecosystem levels, but no clear framework to understand effects of temperature.

**Aim: Develop a eco-physiological model to predict effects of changing temperatures on organism size over ecological and evolutionary time-scales**

# {.col-2}

## Model

We set up a weight ($w$) based model consisting of mass and oxygen balance equations:

Feeding - type 2 response with parameters for maximum consumption rate $\gamma$ and maximum intake $h$, modulated by feeding activity $\tau$:
$$ f = \frac{{\color{uclagold}\tau} \gamma w^{p}\Theta}{{\color{uclagold}\tau}\gamma w^{p}\Theta  + {\color{bondiblue} T^c_{h} h w^{q}}} $$

Mass balance: Intake is assimilated with loss from heat increment $\phi$ and absorbtion $\beta$. Metabolic loss due to standard and feeding activity ($\tau$) metabolic costs
$$ P =(1-\beta-\phi)f {\color{bondiblue} T^c_{h} h w^{q}}-{\color{sinopia}k(1+{\color{uclagold}\tau\delta}) T^c_{M} w^n} $$

Oxygen balance: Oxygen sets the limit for metabolism via the cost of metabolising one gram of feed $\omega$ relative to the potential oxygen supply $f_{O_2}w^n$:
$$P_{0_2} = f_{O_2}w^n - \omega(\beta f_C {\color{bondiblue} T^c_{h} h w^{q}} +{\color{sinopia}k(1+{\color{uclagold}\tau\delta}) T^c_{M} w^n})$$

Mortality: Weight based with base mortality $m_0$ and activity related mortality $m_\tau$.

$$M = (m_0 + {\color{uclagold}\tau}m_{\color{uclagold}\tau})w^{1-p}$$
${\bf \tau}$ **is optimised at any temperature via Gilliam's rule (argmax${\bf _\tau\{P/M\}}$)**

## Ecological dynamics
With increasing temperature, most organisms will increase activity $\tau$ to offset increased metabolic costs.

Feeding level, available energy and mortality are impacted by response to temperature

Oxygen sets metabolic limits at extreme temperatures

```{r}

require(dplyr)
source('../model.R')

# Sluggish small
ss <- list(beta=0.15,
           delta=4,
           Ea=0.52, 
           gamma=40, 
           h=20, 
           k=1.2,
           lO=1, 
           m=100, 
           M=0.1 ,
           n=0.88, 
           O2crit=2, 
           omega=0.25, 
           p=0.8,
           P50=4.2, 
           phi=0.25, 
           q=0.8 ,
           shape=3,
           temp=seq(5,26,l=100),
           Topt=20, 
           temp_ref=15,
           slope=-0.2,
           tmax=30,
           tr=0.3,
           v=6,
           lm=5000,
           nu=-0.2
)

# sluggish large
sl <- ss
sl$nu <- -0.3

# fast large
fl <- sl
fl$lO <- 2
fl$k <- 1.5
fl$h <- 40
fl$v <- 1
fl$M <- 0.3

# fast large
fs <- fl
fs$nu <- -0.2

scenarios <- bind_cols(data_frame(Scenario = as.factor(rep(c('Sluggish small','Sluggish large','Active small','Active large'),each=100))),
                       bind_rows(plot_data_temp(ss),
                                 plot_data_temp(sl),
                                 plot_data_temp(fs),
                                 plot_data_temp(fl)))

ggp <-lst()

```

```{r dev.args= list(bg = 'transparent'),fig.width=7,fig.height=2}

require(ggplot2)
require(cowplot)

sc_act <- tidyr::gather(scenarios,Activity,value,Optimum, Limit,Realised) %>%
  mutate(Activity = relevel(as.factor(Activity), "Realised")) %>%
  filter(!grepl('large',Scenario))

ggplot(sc_act) + 
  geom_line(aes(x=Temperature, y=value, linetype=Activity),alpha=0.7) +
  theme_cowplot()+
  facet_wrap(~Scenario)+
  theme(strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression('Activity ' (tau)))

```
\newline
```{r dev.args= list(bg = 'transparent'),fig.width=7,fig.height=2}

sc <- function(x) x/max(x,na.rm=T)

sc_feed <- scenarios %>%
  filter(!grepl('large',Scenario)) %>% 
  group_by(Scenario) %>%
  mutate(Consumption=sc(Consumption),
         Efficiency=sc(Efficiency),
         Feeding=`Feeding level`,
         `Available Energy` = sc(`C for growth`),
         `Energy` = ifelse(`Available Energy`<0,0,`Available Energy`)) %>% 
  tidyr::gather(Rate,value,Feeding,Energy,M,Efficiency) %>%
  mutate(Rate = relevel(as.factor(Rate),'Feeding'))

ggplot(sc_feed) + 
  geom_line(aes(x=Temperature, y=value, linetype=Rate),alpha=0.7) +
  theme_cowplot()+
  facet_wrap(~Scenario)+
  theme(strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank())+
  scale_linetype_discrete('')+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative rate')

```

Temperature affects size via evolved reaction norm on ecological time-scales:

```{r dev.args= list(bg = 'transparent'),fig.width=7,fig.height=2}

gout_fs <- plot_data_growth_tO2(fs,n_int = 50)

#browser()
norm <- gout_fs$alloc %>% 
  group_by(Temperature) %>%
  summarise(ts=t[ifelse(any(abs(allocs-0.5)<0.1),which.min(abs(allocs-0.5)),NA)-1],
            m=ls[ifelse(any(abs(allocs-0.5)<0.1),which.min(abs(allocs-0.5)),NA)-1])

cp <- ggplot() +
  geom_line(aes(x=t, y=ls,col=as.factor(Temperature)),data=gout_fs$alloc) +
  geom_point(aes(x=ts, y=m),data=norm) +
  #geom_point(data=growth_scenarios[seq.int(1,nrow(growth_scenarios),by = 5),],aes(x=Temperature, y=Temp, col=Scenario,shape=Scenario),alpha=0.5,size=2) +
  #theme_cowplot()+
  viridis::scale_colour_viridis(discrete = T,guide='none')+
  theme(
    panel.background = element_blank(),
    plot.background = element_blank())+
  coord_cartesian()+
  labs(y = 'Length (cm)',
       x='Age  (years)')

scs <- list(ss,sl,fs,fl)
scn <- unique(scenarios$Scenario)
gouts <- lapply(scs,plot_data_growth_tO2,n_int = 50)

alloc <- bind_rows(lapply(1:length(gouts), function(x) {
  y <- gouts[[x]][['alloc']]
  y$Scenario <- scn[x]
  y}))

norm <- alloc %>% 
  group_by(Scenario,Temperature) %>%
  summarise(ts=t[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1],
            ls=ls[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1],
            alloc=allocs[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1]) %>%
  mutate(ls = sc(ls))

np <-  ggplot() +
  geom_line(aes(x=Temperature, y=ls, linetype=Scenario),data=norm) +
  theme(
    panel.background = element_blank(),
    plot.background = element_blank())+
  #scale_color_discrete(guide='none')+
  #scale_shape_discrete(guide='none')+
  coord_cartesian()+
  labs(y = 'Relative length',
       x='Temperature')

cowplot::plot_grid(cp,np,rel_widths = c(0.4,0.6))

```

# {.col-2}

## Traits

Two strategists: minimising mortality $M$ or maximising energy gain $P$

$P$ strategy: 
* Maximise intake at the cost of higher standard metabolism to feeding maintain activity and larger gut.
* Live with higher $M$

$M$ strategy: 
* Minimise $M$ by minimising dangerous feeding activty, minimise resting metabolism

## Evolutionary impact

Size-at maturation reacts strongly to temperature changes in the model, however, changes are small on ecological time-scales ($\lt$1mm per year)

```{r dev.args= list(bg = 'transparent'),fig.width=7,fig.height=3}

winfs <- bind_rows(lapply(1:length(gouts), function(x) {
  y <- gouts[[x]][['winfs']]
  y$Scenario <- scn[x]
  y}))

wtfs <- winfs %>% group_by(Scenario) %>%
  mutate(length=sc(lw(Temp)))

cpG <- ggplot(wtfs) +
  geom_line(aes(x=Temperature, y=length, linetype=Scenario),alpha=0.7) +
  #geom_point(data=growth_scenarios[seq.int(1,nrow(growth_scenarios),by = 5),],aes(x=Temperature, y=Temp, col=Scenario,shape=Scenario),alpha=0.5,size=2) +
  #theme_cowplot()+
  scale_linetype_discrete(guide='none')+
  theme(
    panel.background = element_blank(),
    plot.background = element_blank())+
  #scale_color_discrete(guide='none')+
  #scale_shape_discrete(guide='none')+
  #scale_y_log10()+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative adult length')

npG <- ggplot(wtfs) +
  geom_line(aes(x=Temperature, y=L/length, linetype=Scenario),alpha=0.7) +
  #geom_point(data=growth_scenarios[seq.int(1,nrow(growth_scenarios),by = 5),],aes(x=Temperature, y=Temp, col=Scenario,shape=Scenario),alpha=0.5,size=2) +
  #theme_cowplot()+
  theme(
    panel.background = element_blank(),
    plot.background = element_blank())+
  #scale_color_discrete(guide='none')+
  #scale_shape_discrete(guide='none')+
  coord_cartesian(ylim=quantile(wtfs$L/wtfs$length,c(0.05,0.95),na.rm=T)) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression(Percent~change~y^-1))

cowplot::plot_grid(cpG,npG,rel_widths = c(0.4,0.6))

```

# {.col-1}

## Acknowledgements

P. Neubauer acknowledges funding from the Royal Society of New Zealand Marsden Fast Start grant 14-DFG-001.

All code for this poster is accessble from github.

# {.col-1}
## References