### 2017-09-11 model_meeting.tex

\documentclass[xcolor={table}]{beamer}
\usepackage{multirow} 
\usepackage{caption}
\usepackage{subcaption}
\usepackage{color}
\usetheme{Dragonfly}

\title{Sizing the effects of temparture on fishes}
\subtitle{Current state and future direction}

\projectcode{}
\author{Philipp Neubauer}
\date{\today}

\newcommand{\p}{\bigskip\par}

\begin{document}

%% --------------------------------

\titleslide

\begin{frame}{Aims}

  {\bf Draw a line in the sand - current state \& future
    direction under IEMRS}
  \p

  Current project (PAU2016-03) supports moving data logger project towards
    complete coverage using VMS or electronic reporting
  \p

  IEMRS will require electronic logging on all boats - will
    force complete coverage of electronic reporting on PCELR scale

\end{frame}

\end{document}