### 2017-09-11 poster/poster.Rmd

---
title: "Sizing the effects of temperature on fish"

author: ['<span class=\"presenter\">Philipp Neubauer<sup>1</sup></span>, Ken H. Andersen<sup>2</sup><ol class=\"affiliations\"><li> Dragonfly Data Science, Wellington, New Zealand</li><li>Centre for Ocean Life, National Institute of Aquatic Resources, Technical University of Denmark, Denmark</li></ol>']

output: drposter::revealjs_poster

bibliography: ../Sizing-fx-of-temperature.bib
csl: nature.csl
---

```{r setup, include=FALSE}

require(ggplot2)
require(cowplot)

knitr::opts_chunk$set(echo = FALSE,warning = F,message = F,
                      cache=T,autodep = T,dev = "svg",dev.args= list(bg  = 'transparent'))  # Hide code by default. Can override chunk-by-chunk
```

```{r}

require(dplyr)
source('../model.R')

int_m <- 1000
int_temp <- 40

# Sluggish small
ss <- list(c=0.3,
           beta=0.15,
           delta=4,
           Ea=0.52, 
           gamma=40, 
           h=20, 
           k=1.2,
           lO=1, 
           m=100, 
           M=0.1 ,
           n=0.88, 
           O2crit=2, 
           omega=0.25, 
           p=0.8,
           P50=4.2, 
           phi=0.25, 
           q=0.8 ,
           shape=3,
           temp=seq(5,26,l=int_temp),
           Topt=20, 
           temp_ref=15,
           slope=-0.2,
           tmax=30,
           tr=0.9,
           v=6,
           lm=int_m,
           nu=-0.2
)

# fast small
fs <- ss
fs$lO <- 2
fs$k <- 1.5
fs$h <- 40
fs$v <- 6
fs$M <- 0.3
fs$tmax <- 20
fs$nu <- -0.2

scenarios <- bind_cols(data_frame(Scenario = as.factor(rep(c('M strategy','P strategy'),each=int_temp))),
                       bind_rows(plot_data_temp(ss),
                                 plot_data_temp(fs)))

ggp <-lst()

```

# {.col-1}

## Introduction

&nbsp;

Size-at-maturation and maximum size are key traits, defining ecological niches and driving ecological interactions [@andersen_asymptotic_2006]. Temperature has been identified in various contexts as a key driver of organism size (Bergman's rule, Temperature-size-rule...) [@atkinson_temperature_1994]. Climate change may therefore have strong effects on organism size and follow on effects at population and ecosystem levels, but no clear framework to understand effects of temperature [@lefevre_models_2017].

&nbsp;

**Aim: Develop a eco-physiological model [@holt_climate_2014] to predict effects of changing temperatures on organism size over ecological and evolutionary time-scales**

# {.col-2}

## Model

&nbsp;

We set up a weight ($w$) based model consisting of mass and oxygen balance equations:

Feeding - type 2 functional response with parameters for maximum consumption rate $\gamma$, modulated by feeding activity $\tau$, and maximum intake $h$, which responds to temperature via $T^c_{M}$:
$$ f = \frac{{\color{uclagold}\tau} \gamma w^{p}\Theta}{{\color{uclagold}\tau}\gamma w^{p}\Theta  + {\color{bondiblue} T^c_{h} h w^{q}}} $$

Mass balance: Intake is assimilated with loss from heat increment $\phi$ and absorption $\beta$. Metabolic loss due to standard and feeding activity ($\tau$) metabolic costs, modulated by temperature via $T^c_{M}$
$$ P =(1-\beta-\phi)f {\color{bondiblue} T^c_{h} h w^{q}}-{\color{sinopia}k(1+{\color{uclagold}\tau\delta}) T^c_{M} w^n} $$

Oxygen balance: Oxygen sets the limit for metabolism via the cost of metabolising one gram of feed $\omega$ relative to the potential oxygen supply $f_{O_2}w^n$:
$$P_{0_2} = f_{O_2}w^n - \omega(\beta f_C {\color{bondiblue} T^c_{h} h w^{q}} +{\color{sinopia}k(1+{\color{uclagold}\tau\delta}) T^c_{M} w^n})$$

Mortality: Weight based with base mortality $m_0$ and activity related mortality $m_\tau$.

$$M = (m_0 + {\color{uclagold}\tau}m_{\color{uclagold}\tau})w^{1-p}$$
${\bf \tau}$ **is optimised at any temperature via Gilliam's rule (argmax${\bf _\tau\{P/M\}}$)**

&nbsp;

Strategy trade-off: minimising mortality $M$ or maximising energy gain $P$

&nbsp;

$P$ strategy: Maximise intake at the cost of higher standard metabolism to maintain feeding activity and larger gut. Live with higher $M$.

&nbsp;

$M$ strategy: Minimise $M$ by minimising dangerous feeding activity, minimise resting metabolism.

## Temperature impact on size

&nbsp;

```{r fig.width=7,fig.height=2.5,fig.cap='Figure 3: **On ecological time-scales**, temperature affects size via evolved maturation-reaction norms [@marty_impact_2011]. Most reaction norms found in nature have a negative slope, meaning for faster growth, maturation occurs at a larger size. Increasing temparture (left panel, purple to yellow growth curves) modifies growth, and maturation age changes according to the reaction norm (right panel).'}

gout_fs <- plot_data_growth_tO2(fs)

#browser()
norm <- gout_fs$t_growth %>% 
  group_by(Temperature) %>%
  summarise(ts=t[ifelse(any(abs(allocs-0.5)<0.1),which.min(abs(allocs-0.5)),NA)-1],
            m=t_length[ifelse(any(abs(allocs-0.5)<0.1),which.min(abs(allocs-0.5)),NA)-1])

cp <- ggplot() +
  geom_line(aes(x=t, y=t_length,col=as.factor(Temperature)),data=gout_fs$t_growth) +
  geom_point(aes(x=ts, y=m),data=norm) +
  #geom_point(data=growth_scenarios[seq.int(1,nrow(growth_scenarios),by = 5),],aes(x=Temperature, y=Temp, col=Scenario,shape=Scenario),alpha=0.5,size=2) +
  theme_cowplot()+
  viridis::scale_colour_viridis(discrete = T,guide='none')+
  theme(plot.margin = margin(2, 6, 0, 6),
        panel.background = element_blank(),
        plot.background = element_blank())+
  coord_cartesian()+
  labs(y = 'Length (cm)',
       x='Age  (years)')

scs <- list(ss,fs)
scn <- unique(scenarios$Scenario)
gouts <- parallel::mclapply(scs,plot_data_growth_tO2,mc.cores = 4)

alloc <- bind_rows(lapply(1:length(gouts), function(x) {
  y <- gouts[[x]][['t_growth']]
  y$Scenario <- scn[x]
  y}))

norm <- alloc %>% 
  group_by(Scenario,Temperature) %>%
  summarise(ts=t[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1],
            ls=t_length[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1],
            alloc=allocs[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1]) %>%
  mutate(ls = sc(ls))

np <-  ggplot() +
  geom_line(aes(x=Temperature, y=ls, linetype=Scenario),data=norm) +
  theme_cowplot()+
  theme(plot.margin = margin(2, 6, 0, 6),
        panel.background = element_blank(),
        plot.background = element_blank())+
  #scale_color_discrete(guide='none')+
  #scale_shape_discrete(guide='none')+
  coord_cartesian()+
  labs(y = 'Relative length',
       x='Temperature')

cowplot::plot_grid(cp,np,rel_widths = c(0.4,0.6))

```

&nbsp;

```{r fig.width=7,fig.height=2.5,fig.cap='Figure 4: **On evolutionary time-scales**, selection moves the reaction norm. Size-at-maturation reacts strongly to temperature changes in the model (left panel). Selection (S) induced changes in size are slow for M strategists, fast for P strategists.'}

winfs <- bind_rows(lapply(1:length(gouts), function(x) {
  y <- gouts[[x]][['winfs']]
  y$Scenario <- scn[x]
  y}))

wtfs <- winfs %>% group_by(Scenario) %>%
  mutate(length=sc(lw(l)))

cpG <- ggplot(wtfs) +
  geom_line(aes(x=Temperature, y=length, linetype=Scenario),alpha=0.7) +
  theme_cowplot()+
  scale_linetype_discrete(guide='none')+
  theme(plot.margin = margin(3, 6, 0, 6),
        panel.background = element_blank(),
        plot.background = element_blank())+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative adult length')

npG <- ggplot(wtfs) +
  geom_line(aes(x=Temperature, y=L/length, linetype=Scenario),alpha=0.7) +
  theme(plot.margin = margin(3, 6, 0, 6),
        panel.background = element_blank(),
        plot.background = element_blank())+
  coord_cartesian(ylim=quantile(wtfs$L/wtfs$length,c(0.05,0.95),na.rm=T)) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression(S~response~y^-1))

cowplot::plot_grid(cpG,npG,rel_widths = c(0.4,0.6))

```

# {.col-2}

## Ecological dynamics

```{r fig.width=7,fig.height=2,fig.cap='Figure 1: With increasing temperature, most organisms will increase activity $\\tau$ to offset increased metabolic costs. Oxygen sets metabolic limits at extreme temperatures.'}

sc_act <- tidyr::gather(scenarios,Activity,value,Optimum, Limit,Realised) %>%
  mutate(Activity = relevel(as.factor(Activity), "Realised")) %>%
  filter(!grepl('large',Scenario))

ggplot(sc_act) + 
  geom_line(aes(x=Temperature, y=value, linetype=Activity),alpha=0.7) +
  theme_cowplot()+
  facet_wrap(~Scenario)+
  theme(plot.margin = margin(0, 6, 0, 6),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression('Activity ' (tau)))

```

```{r fig.width=7,fig.height=2,fig.cap='Figure 2: Feeding level, available energy and mortality are impacted by increased activity and the metabolic response to temperature.'}

sc_feed <- scenarios %>%
  filter(!grepl('large',Scenario)) %>% 
  group_by(Scenario) %>%
  mutate(Consumption=sc(Consumption),
         Efficiency=sc(Efficiency),
         Feeding=`Feeding level`,
         `Available Energy` = sc(`C for growth`),
         `Energy` = ifelse(`Available Energy`<0,0,`Available Energy`)) %>% 
  tidyr::gather(Rate,value,Feeding,Energy,M,Efficiency) %>%
  mutate(Rate = relevel(as.factor(Rate),'Feeding'))

ggplot(sc_feed) + 
  geom_line(aes(x=Temperature, y=value, linetype=Rate),alpha=0.7) +
  theme_cowplot()+
  facet_wrap(~Scenario)+
  theme(plot.margin = margin(0, 6, 0, 6),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank())+
  scale_linetype_discrete('')+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative rate')

```

## Take home messages

<span style="font-size: 300%;">

&nbsp;

&nbsp;

* **On ecological time-scales changes in adult size due to temperature alone will likely be small, and masked by changes in the environment.**

&nbsp;

&nbsp;

* **On evolutionary time-scales, size responds strongly. The relative selection response is strongest at temperature extremes.**

&nbsp;

&nbsp;

* **Size of $P$ strategy species with high activity reacts more strongly compared to sluggish $M$ strategy species.**

</span>

# {.col-1}

## Acknowledgements & References

P. Neubauer acknowledges funding from the Royal Society of New Zealand Marsden Fast Start grant 14-DFG-001. All code for the model and this poster is accessible from github at repository Adaptive-dynamics/Sizing-fx-of-temperature.

&nbsp;