### 2018-07-06 figures/O2_fig.Rnw

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(warning=F, message = FALSE,echo=F,results='asis',error=FALSE,fig.lp='fig:',fig.path='images/')
@

<<O2_fig,fig.cap='Maximum metabolic rate, determined by oxygen supply, as a function of temperature, for four different optimal temperatures (panels---see panel label for optimal temperature in degrees celsius), and increasing values of $\\eta$.', fig.width=8, fig.height=4, echo=FALSE, out.width='\\linewidth'>>=

grid <- expand.grid(eta = c(0.1,0.5,3,5),
                    Topt = c(10,15,20,25))

O2fun <- function(x) {
  ts = with(x,((26-seq(1,26,l=1000))/(26-Topt))^eta*exp(-eta*(26-seq(1,26,l=1000))/(26-Topt)))
  ts = ts/max(ts)
}

require(dplyr)
require(purrr)
require(ggplot2)
#require(cowplot)

g <- grid %>% 
  mutate(n=1:n()) %>%
  split(.$n) %>%
  map(~data.frame(eta = .$eta,
                  Topt = .$Topt,
                  O2 = O2fun(.))) %>%
  bind_rows(.id = 'n') %>%
  group_by(eta,Topt) %>%
  mutate(Temperature=seq(1,26,l=1000)) 
# 
# ggplot(g) + 
#   geom_line(aes(x = Temperature, 
#                 y = O2,
#                 linetype=factor(eta))) + 
#   theme_bw() + 
#   facet_grid(Topt~.) +
#   scale_linetype_discrete(bquote(eta)) + 
#   scale_x_continuous(limits = c(3,29),breaks = seq(4,28,4)) +
#   labs(y=expression(Relative~O[2]~supply),
#        x=expression("Temperature " ( degree*C)))

@

<<O2_fig_used,fig.cap='Maximum oxygen supply relative to the maximum supply for species with a dome-shaped MOS (here $\\eta$ = 3) and a continually increasing MOS ($\\eta$ = 0.1) used in model scenarios discussed below.', fig.width=6, fig.height=3, echo=FALSE, out.width='\\linewidth'>>=

ggplot(g %>% filter((eta==0.1 & Topt==25)|(eta==3 & Topt==20))) + 
  geom_line(aes(x = Temperature, 
                y = O2,
                linetype=factor(eta))) + 
  theme_bw() + 
  scale_linetype_discrete(bquote(eta)) + 
  scale_x_continuous(limits = c(3,29),breaks = seq(4,28,4)) +
  labs(y=expression(Relative~O[2]~supply),
       x=expression("Temperature " ( degree*C)))
  #scale_color_discrete(bquote(T[opt]), h.start = 180)

@